---
title: "SpringsFire Data Analysis"
author: "Ashley Grupenhoff"
date: "4/19/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---



# Prescribed Monitoring Field Crew Initial Analysis 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(tidyverse)
  library(tinytex)
          options(tinytex.verbose=TRUE)
library(tibble)
library(dplyr)
library(ggplot2)
library(vegan)
library(knitr)
library(kableExtra)
library(rio)
      
```

## 2020 post Springs fire fuels

Initial fuels data 

0-3" = (11.64 * n * d^2 * s * a * c)/ (Nl)
CWD = (11.64 * n * sum(d^2) * s * a * c)/ (Nl)

slope correction = (1+(presprings_fuelslope/100)^2)^(1/2)
convert sq cm to sq in: * 0.155

```{r}
#load data

post_springs_20_cwd <- read_csv("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/raw/2020/post_springsfire_CWD_20.csv")
post_springs_20_fuels <- read_csv("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/raw/2020/post_springsfire_finefuels_20.csv")
SpringsFire_treatment <- read_csv("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/raw/springsfire_treatment.csv")

#CWD calculations
post_springs_20_cwd <- na.omit(post_springs_20_cwd) 

post_springs_20_cwd <- post_springs_20_cwd %>% 
                            mutate( decay_type = case_when(
                            decay >= 4 | decay == '5-rotten' ~ "rotten",
                            decay < 4 ~ "sound"
                          )) %>% 
                      mutate(diam_in_square = (intersect_cm * 0.3937)^2) %>% 
                      group_by(plotid, decay_type) %>% 
                      summarise(sum_diam = sum(diam_in_square)) %>% 
                      pivot_wider(names_from = "decay_type", values_from = "sum_diam") 

#replaceNAs with 0

post_springs_20_cwd[is.na(post_springs_20_cwd)]=0
                      

                      
             
```


```{r, message=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results= "asis"}

#Create New data frame
                    postspringsfuels20<-NULL
                    postspringsfuels20<-as.data.frame(matrix(NA, ncol=6, nrow=nrow(PostSprings20Fuels_Slope)))
                    colnames(postspringsfuels20)<-c("plot_id","mass_1hr", "mass_10hr", "mass_100hr", "mass_cwd_sound", "mass_cwd_rotten")
                    
                    #run a forloop to apply the equations to each row in the data frame
                    for (i in 1:nrow(PostSprings20Fuels_Slope)) {
                    	      
                      postspringsfuels20[i,"plot_id"]<-postsprings20_fuels_combined[i,"plot_id"]          
                      postspringsfuels20[i,"mass_1hr"]<- ((11.64 * postsprings20_fuels_combined[i,"count_x1h"] * 0.0151 * 	0.48 * 1.13 * postsprings20_fuels_combined [i, "slopecorrection"])/ (2*3.28))                          
                      postspringsfuels20[i,"mass_10hr"]<-((11.64 * postsprings20_fuels_combined[i,"count_x10h"] * 0.289 * 	0.48 * 1.13 * postsprings20_fuels_combined[i, "slopecorrection"]) / (2*3.28))                            
                      postspringsfuels20[i,"mass_100hr"]<-((11.64 * postsprings20_fuels_combined[i,"count_x100h"] * 2.76 * 	0.40 * 1.13 * postsprings20_fuels_combined[i, "slopecorrection"]) / (4*3.28))           
                      postspringsfuels20[i,"mass_cwd_sound"]<- ((11.64 * postsprings20_fuels_combined[i,"sound"] 	* 0.155 * 0.40 * postsprings20_fuels_combined[i, "slopecorrection"]) / (11.3*3.28))           
                      postspringsfuels20[i,"mass_cwd_rotten"]<-((11.64 * postsprings20_fuels_combined[i,"rotten"] 	* 0.155 * 0.30 * postsprings20_fuels_combined[i, "slopecorrection"]) / (11.3*3.28))    
                            
                    }      
                       
##Average each mass type per plot
              
              postspringsfuels20_complete<- postspringsfuels20 %>% 
                            mutate_all(~replace(., is.na(.), 0)) %>% 
                            group_by(plot_id) %>% 
                            mutate(total_fuel = sum(mass_1hr,mass_10hr,mass_100hr,mass_cwd_sound, mass_cwd_rotten)) %>% 
                              pivot_longer(cols = -plot_id, names_to = "fuel_size", values_to= "mass") 
                            

                        
                        #add column for treatment type
                        
                        postspringsfuels20_complete <- left_join(postspringsfuels20_complete,SpringsFire_treatment,                                                                                            by = "plot_id")
                        
                       

                      #save completed post fuels data frame
                      postspringsfuels20_complete <-   import("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/postspringsfuels20_complete.csv")



```


## pre and post burn fuel consumption

Comparison of fuels consumption by treatment (time since last burn). Only includes plots in which we have pre- and post-burn 2019 data. 


```{r, echo=FALSE, message=FALSE, warning=FALSE,results= "asis"}

prespringsfuels_complete_joined$preburn_mass <- prespringsfuels_complete_joined$mass
postspringsfuels_complete_joined$postburn_mass <- postspringsfuels_complete_joined$mass

prespringsfuels_complete_joined <- prespringsfuels_complete_joined %>% 
  mutate(plot_id_fuel = paste(plot_id, fuel_size))
postspringsfuels_complete_joined <- postspringsfuels_complete_joined %>% 
  mutate(plot_id_fuel = paste(plot_id, fuel_size))

springsfirefuels_ALL <- prespringsfuels_complete_joined %>% 
  left_join(postspringsfuels_complete_joined,prespringsfuels_complete_joined, by = "plot_id_fuel") %>% 
  select(plot_id.x, fuel_size.x, preburn_mass, postburn_mass, treatment_type.x) 
  
springsfirefuels_completeplots <- springsfirefuels_ALL %>% 
   na.omit() %>% 
   mutate(id_fuelsize = paste(plot_id.x, fuel_size.x)) %>% 
  select(id_fuelsize, preburn_mass, postburn_mass) %>% 
  pivot_longer(-id_fuelsize, names_to = "time", values_to = "mass") %>% 
  separate(id_fuelsize, c("plot_id", "fuel_size"), sep=" ") 

springsfirefuels_completeplots <- left_join(springsfirefuels_completeplots, SpringsFire_treatment, by="plot_id")

#reorder so preburn comes before postburn
springsfirefuels_completeplots$time <- factor(springsfirefuels_completeplots$time, levels=c("preburn_mass", "postburn_mass"))


export(springsfirefuels_completeplots, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/springsfuels_complete.csv")

springs_fuels_complete_plot <- ggplot(data=springsfirefuels_completeplots, mapping = aes(x=treatment_type, y=mass, fill=time))+
              geom_boxplot()+
              
              facet_wrap(~ fuel_size, scales = "free_y") +
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Mass (Tons/Acre)")+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              theme(axis.text.x = element_text(angle=45, hjust=1))+
             
              theme(plot.title = element_text(hjust = 0.5))
springs_fuels_complete_plot

ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springsfuelscomplete.png", springs_fuels_complete_plot)

springs_fuels_complete2 <- ggplot(data=springsfirefuels_completeplots, mapping = aes(x=time, y=mass, fill=time))+
              geom_boxplot()+
  scale_fill_manual(values=c("#69b3a2", "#404080"))+
              
              facet_wrap(~ fuel_size, scales = "free_y") +
              theme_minimal()+
              
              ylab("Mass (Tons/Acre)")+
              theme(axis.text.x = element_text(angle=45, hjust=1))+
              theme(plot.title = element_text(hjust = 0.5))

springs_fuels_complete2

ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springsfuelscomplete2.png", springs_fuels_complete2)



##LITTER DUFF CONSUMPTION

                Pre_springs_litter_duff <- Pre_SpringsFire_FUELS %>% 
                  select(plot_id, duff_depth_cm, litter_depth_cm) %>% 
                  group_by(plot_id) %>% 
                   summarize(mean_litter = mean(litter_depth_cm),
                         mean_duff = mean(duff_depth_cm)) %>% 
                  pivot_longer(-plot_id, names_to = "litter_duff", values_to = "depth_cm")
                export(Pre_springs_litter_duff, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/pre_springs_litterduff.csv")
                
                  
                Post_springs_litter_duff <- Post_SpringsFire_FUELS %>% 
                  select(plot_id, duff_depth_cm, litter_depth_cm) %>% 
                  group_by(plot_id) %>% 
                   summarize(mean_litter = mean(litter_depth_cm),
                         mean_duff = mean(duff_depth_cm)) %>% 
                    na.omit() %>% 
                  pivot_longer(-plot_id, names_to = "litter_duff", values_to = "depth_cm")
                export(Post_springs_litter_duff, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/post_springs_litterduff.csv")


                Post_springs_litter_duff <- Post_springs_litter_duff %>% 
                  mutate(plot_id_litterduff = paste(plot_id, litter_duff))
                Pre_springs_litter_duff <- Pre_springs_litter_duff %>% 
                  mutate(plot_id_litterduff = paste(plot_id, litter_duff))
                
                springsfire_litterduffall <-   left_join(Pre_springs_litter_duff, Post_springs_litter_duff, by = "plot_id_litterduff") 
                springsfire_litterduffall$preburn_depth_cm <- springsfire_litterduffall$depth_cm.x 
                springsfire_litterduffall$postburn_depth_cm <- springsfire_litterduffall$depth_cm.y
                
                springsfire_litterduffall <-  springsfire_litterduffall %>% 
                      select(plot_id_litterduff, preburn_depth_cm, postburn_depth_cm) %>% 
                  na.omit() %>% 
                   pivot_longer(-plot_id_litterduff, names_to = "time", values_to = "depth_cm") 
                
                springsfire_litterduffall <- springsfire_litterduffall %>% 
                 separate(plot_id_litterduff, c("plot_id", "litter_duff"), sep=" ")
                

#MAKE PRE BURN FIRST
          springsfire_litterduffall$time <- factor(springsfire_litterduffall$time, levels=c("preburn_depth_cm", "postburn_depth_cm"))

str(springsfire_litterduffall$time)

export(springsfire_litterduffall, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/springsfire_litterduff_consumption.csv")

springs_litter_duff <- ggplot(data=springsfire_litterduffall, mapping = aes(x=litter_duff, y=depth_cm, fill=time))+
              geom_boxplot()+
             
              theme_minimal()+
              xlab("Duff and Litter consumption")+
              ylab("Depth (cm)")+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              theme(axis.text.x = element_text(angle=45, hjust=1))+
             
              theme(plot.title = element_text(hjust = 0.5))
springs_litter_duff

ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springsfuels_litterduff.png", springs_litter_duff)


springsfuels <- import("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/springsfuels_complete.csv")

springsfuels_total <- springsfuels %>% 
  pivot_wider(names_from = fuel_size, values_from = mass) %>% 
                                          group_by(plot_id,time) %>% 
                          mutate(total_fuel = sum(mass_1hr,mass_10hr,mass_100hr,mass_cwd_sound, mass_cwd_rotten)) %>% 
                          mutate(plot_id_time = paste(plot_id, time)) %>% 
                          group_by(plot_id_time) %>% 
                            select(plot_id_time, mass_1hr, mass_10hr, mass_100hr, mass_cwd_sound, mass_cwd_rotten, total_fuel) %>% 
                              pivot_longer(-plot_id_time, names_to = "fuel_size", values_to = "mass") %>% 
                              separate(plot_id_time, c("plot_id", "time"), sep=" ") 
springsfuels_total <- left_join(springsfuels_total, SpringsFire_treatment, by="plot_id")


  
```


## Tree data

```{r, echo=FALSE, message=FALSE, warning=FALSE, results= "asis"}
Pre_SpringsFire_trees <- read_csv("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/raw/pre_springsfire_trees.csv", col_names = TRUE)

Pre_SpringsFire_trees$ht_to_crownlive_m <- as.numeric(Pre_SpringsFire_trees$ht_to_crownlive_m )


#live trees only


                    presprings_trees_live <- Pre_SpringsFire_trees %>% 
                      filter(status == "live") %>% 
                      select(plot_id, species_id, dbm_cm, ht_m, ht_to_crownlive_m) %>% 
                      drop_na() %>% 
                      group_by(plot_id, species_id) %>% 
                      summarize(mean_dbh = mean(dbm_cm), 
                                mean_ht = mean(ht_m),
                                mean_ht2crwn = mean(ht_to_crownlive_m),
                                total_trees = length(dbm_cm)) %>% 
                      mutate(total_trees_acre = total_trees*10) 



                          presprings_trees_live <-  left_join(presprings_trees_live, SpringsFire_treatment, by="plot_id")
                          
                          presprings_trees_live <- drop_na(presprings_trees_live)
                          
                          export(presprings_trees_live, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/prespringstrees_complete.csv")


  #average density per plot
                    
                          #live      
                     presprings_trees_live_total <- presprings_trees_live %>% 
                       group_by(plot_id) %>% 
                       summarize(total_trees_acre = sum(total_trees_acre))
                     
                     mean(presprings_trees_live_total$total_trees_acre)
                          
                          #adjust values to be per acre


springstotaltrees_plot <- ggplot(data=presprings_trees_live, mapping = aes(x=treatment_type, y=total_trees_acre, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7)+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Total adult live trees per acre (count)")+
   ggtitle("Total Live Trees per Acre")+
              theme(plot.title = element_text(hjust = 0.5))

springs_avgdbh_plot <- ggplot(data=presprings_trees_live, mapping = aes(x=treatment_type, y=mean_dbh, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7)+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Average DBH (cm)")+
     ggtitle("Average DBH (cm) Live Trees")+
              theme(plot.title = element_text(hjust = 0.5))

springs_avght_plot <- ggplot(data=presprings_trees_live, mapping = aes(x=treatment_type, y=mean_ht, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7)+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Mean Height (m)")+
  ggtitle("Average Height (m) Live Trees")+
              theme(plot.title = element_text(hjust = 0.5))

springs_avght2crwn_plot <- ggplot(data=presprings_trees_live, mapping = aes(x=treatment_type, y=mean_ht2crwn, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7)+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Mean Height to Live Crown (m)")+
  ggtitle("Average Height to Crown (m) Live Trees")+
              theme(plot.title = element_text(hjust = 0.5))

ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springstotaltrees_plot.png", springstotaltrees_plot)
ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springs_avgdbh_plot.png", springs_avgdbh_plot)
ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springs_avght_plot.png", springs_avght_plot)
ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springs_avght2crwn_plot.png", springs_avght2crwn_plot)






#live and dead trees simple comparison


presprings_trees_standing <- Pre_SpringsFire_trees %>% 
  filter(status == "live" | status == "dead") %>% 
  select(plot_id, status, species_id, dbm_cm, ht_m) %>% 
  group_by(plot_id, species_id, status) %>% 
  summarize(mean_dbh = mean(dbm_cm), 
            mean_ht = mean(ht_m),
           
            total_trees = length(dbm_cm)) %>% 
  mutate(total_trees_acre = total_trees*10) 
presprings_trees_standing <-  left_join(presprings_trees_standing, SpringsFire_treatment, by="plot_id")
presprings_trees_standing <- drop_na(presprings_trees_standing)
export(presprings_trees_standing, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/prespringstrees_standing.csv")

ggplot(data=presprings_trees_standing, mapping = aes(x=treatment_type, y=total_trees_acre, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              facet_wrap(~status,scales = "free_y")+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7)+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Total adult trees per acre (count)")+
  ggtitle("Total Standing Trees per Acre")+
              theme(plot.title = element_text(hjust = 0.5))

ggplot(data=presprings_trees_standing, mapping = aes(x=treatment_type, y=mean_dbh, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              facet_wrap(~status,scales = "free_y")+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7)+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Mean DBH (cm)")+
  ggtitle("Average DBH (cm) Standing Trees")+
              theme(plot.title = element_text(hjust = 0.5))

ggplot(data=presprings_trees_standing, mapping = aes(x=treatment_type, y=mean_ht, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
   facet_wrap(~status,scales = "free_y")+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7)+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Mean Height (m)")+
  ggtitle("Average Height Standing Trees")+
              theme(plot.title = element_text(hjust = 0.5))

DT::datatable(presprings_trees_standing, options = list(pageLength = 10))
  

```



>Additional data that I have not provided: look at only dead trees, look at stumps, and health/decay by trt.

## Initial basal area

Basal area compared between treatment type, by species and status (live or dead).

*Note* This includes a chunk of control plots that were PICO dominated. Might be best to take these out for statistical analysis. Just plotting all data for now. 

```{r, echo=FALSE, message=FALSE, warning=FALSE, results= "asis"}
Pre_SpringsFire_ba <- read_csv("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/raw/pre_springsfire_basalarea.csv", col_names = TRUE)

Pre_SpringsFire_ba <- na.omit(Pre_SpringsFire_ba)
Pre_SpringsFire_ba <-   left_join(Pre_SpringsFire_ba, SpringsFire_treatment, by="plot_id") 

SpringsFire_ba_condense <- Pre_SpringsFire_ba %>% 
    select(plot_id, status, species_id, stand_ba, treatment_type) 


SpringsFire_ba_condense$stand_ba <- as.numeric(SpringsFire_ba_condense$stand_ba)

export(SpringsFire_ba_condense, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/prespringsbasalarea_complete.csv")

springs_basalarea <- ggplot(data=SpringsFire_ba_condense, mapping = aes(x=treatment_type, y=stand_ba, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7,aes(colour = status))+
              scale_color_manual(values=c("#69b3a2", "#404080"))+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Initial stand basal area (ft^2/acre)")+
  ggtitle("Stand Basal Area (ft^2/acre)")+
              theme(plot.title = element_text(hjust = 0.5))
springs_basalarea
              
mean(SpringsFire_ba_condense$stand_ba)

ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springs_basalarea.png", springs_basalarea)
  
```



## Initial Regen


```{r, echo=FALSE, message=FALSE, warning=FALSE, results= "asis"}
Pre_SpringsFire_regen <- read_csv("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/raw/pre_springsfire_regen.csv")

Pre_SpringsFire_regen <-  left_join(Pre_SpringsFire_regen, SpringsFire_treatment, by="plot_id")

Pre_SpringsFire_regen$tse_count <- as.numeric(Pre_SpringsFire_regen$tse_count)


#Seedling Count

springsfire_regen_tse <- Pre_SpringsFire_regen %>% 
  filter(lifeform == "TSE") %>% 
select(plot_id, species_id, status, lifeform, tse_count, tse_ht_cm) %>% 
   group_by(plot_id, species_id, status) %>% 
  summarize(total_tse_count = sum(tse_count),
            avg_tse_ht = mean(tse_ht_cm)) %>% 
  mutate(tse_count_acre = total_tse_count*10)

springsfire_regen_tse <- left_join(springsfire_regen_tse, SpringsFire_treatment, by="plot_id")
springsfire_regen_tse <- drop_na(springsfire_regen_tse)




springs_tseplot <- ggplot(data=springsfire_regen_tse, mapping = aes(x=treatment_type, y=tse_count_acre, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7,aes(colour = status))+
              scale_color_manual(values=c("#69b3a2", "#404080"))+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Seedling Count (per Acre)")+
  ggtitle("Seedling Count (per Acre)")+
              theme(plot.title = element_text(hjust = 0.5))
ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springs_tseplot.png", springs_tseplot)

DT::datatable(springsfire_regen_tse, options = list(pageLength = 10))



#Sapling Count

springsfire_regen_tsa <- Pre_SpringsFire_regen %>% 
  filter(lifeform == "TSA") %>% 
select(plot_id, species_id, status, lifeform, tsa_dbh_cm, tsa_ht_m, tsa_age_yr) %>% 
  group_by(plot_id, species_id, status) %>% 
  summarize(total_tsa_count = length(tsa_dbh_cm),
            avg_tsa_ht = mean(tsa_ht_m),
            avg_tsa_age = mean(tsa_age_yr)) %>%   
  mutate(tsa_count_acre = total_tsa_count*10)

springsfire_regen_tsa <- left_join(springsfire_regen_tsa, SpringsFire_treatment, by="plot_id")

export(springsfire_regen_tse, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/prespringsregen_tse_complete.csv")
export(springsfire_regen_tsa, "C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/data/clean/prespringsregen_tsa_complete.csv")
  
springs_tsaplot <- ggplot(data=springsfire_regen_tsa, mapping = aes(x=treatment_type, y=tsa_count_acre, fill = species_id))+
              scale_fill_manual(values=c("#69b3a2", "#404080"))+
              geom_boxplot(alpha=5)+
              geom_jitter(alpha= 0.7,aes(colour = status))+
              scale_color_manual(values=c("#69b3a2", "#404080"))+
              theme_minimal()+
              xlab("Prior Treatment")+
              ylab("Sapling Count (per Acre)")+
  ggtitle("Sapling Count (per Acre)")+
              theme(plot.title = element_text(hjust = 0.5))
ggsave("C:/Users/ashle/Documents/R/RX_Fire_Monitoring/RX_Fire_Monitoring/Springs_Fire/images/springs_tsaplot.png", springs_tsaplot)

DT::datatable(springsfire_regen_tsa, options = list(pageLength = 10))

```



