---
title: "Biomass Calculations 2019"
output: html_document
authors: Sara Winsemius and Daniel O'Brien
---

Note to self: all biomass measures are slope corrected. Maybe don't do that? Need to look into it more.

1. Load in data and source functions.
```{r, message=FALSE, warning=FALSE}
library(raster)

source("biomass_equations_2019.R")

# read in tree and cwd data
tree_data <- read.csv("../../data/saved/tree_data_all.csv", stringsAsFactors = FALSE)
cwd_data <- read.csv("../../data/saved/cwd_data_all.csv", stringsAsFactors = FALSE)

tree_data$BA <- pi*(tree_data$DBH_cm)^2/40000
```

1.5. Extract slope data, and aspect while we're at it. NOTE: This chunk will not work because there is no plot data yet. If we have an all-plots data file we can use that, otherwise maybe this can be cut in later on in this script?

```{r, warning=FALSE}
library(raster)
library(sf)
library(rgdal)

# load topo data
slope <- raster("../../data/RS_data/topo/slope_mosaic.tif")
aspect <- raster("../../data/RS_data/topo/aspect_mosaic.tif")
# topo data is in NAD83, same as the GPS unit, so get the CRS for the plot data
nad83 <- st_crs(slope)
# set plot data to spatial_points_dataframe
tree_plot_data <- tree_data[ , 1:7]
cwd_plot_data <- cwd_data[ , 1:7]

all_data <- rbind(tree_plot_data, cwd_plot_data)

all_plots <- summarize(group_by(all_data, PlotID, Date, Observers, FireSev, Easting, Northing, PointAccuracy))

plots <- st_as_sf(all_plots, coords = c("Easting", "Northing"), crs = nad83)

# extract slope and aspect 
stack <- stack(slope, aspect)
topo <- extract(stack, plots)

plots <- cbind(plots, topo)

head(plots)
#which(plots$slope == 72) #72% slope is max
#plots[41,] # PAL303 slope from our plot data is 78%
#plots[111,] # LAK304 slope is 58%

plots$aspect <- plots$aspect_mosaic
plots$slope <- plots$slope_mosaic
del <- c("aspect_mosaic", "slope_mosaic")
plots <- plots[, !(names(plots) %in% del)] # taking out
head(plots)
#st_write(plotshp, "../../plots/biomass_allplots.shp", driver = "ESRI shapefile")

# change object name to maintain consistency in script
slope_data <- data.frame(PlotID = rep(0, nrow(plots)))
slope_data$PlotID <- plots$PlotID
slope_data$aspect <- plots$aspect
slope_data$slope <- plots$slope
```

2. Calculate biomass for each tree and cwd entry and group by plot.

```{r, warning=FALSE}
library(dplyr)
library(purrr)
library(tidyr)

# apply biomass calculation to all tree entries
tree_data$Biomass_kg <- pmap_dbl(list(species = tree_data$Species, dia = tree_data$DBH_cm, tht = tree_data$Height_m), calc_tree_biomass)

# group data by plot and calculate sum of tree biomass
tree_plot_data <- summarize(group_by(tree_data, PlotID, Date, Observers, FireSev, Northing, Easting, PointAccuracy), LiveTreeBiomass_kg = sum(Biomass_kg[which(Status == "L")]), DeadTreeBiomass_kg = sum(Biomass_kg[which(Status == "D")]), BA = sum(BA))

# calculate squared diameter for cwd biomass calculation
cwd_data$DiameterIntersectSquared_cmsq <- cwd_data$DiameterIntersect_cm^2

# assign decay type to each cwd entry
cwd_data$DecayType <- if_else(cwd_data$DecayClass %in% c(1, 2, 3), "Sound", "Rotten")

# join cwd data with slope data
cwd_data <- left_join(cwd_data, slope_data, by = "PlotID")

# calculate slope factor for cwd biomass calculation
cwd_data$SlopeFactor <- sqrt(1 + (cwd_data$slope/100)^2)

# group cwd entries by plot and calculate sum of sound and rotten cwd squared diameters
cwd_plot_data <- summarize(group_by(cwd_data, PlotID, Date, Observers, FireSev, Northing, Easting, PointAccuracy, SlopeFactor), SoundSumDiameterIntersectSquared_cmsq = sum(DiameterIntersectSquared_cmsq[which(DecayType == "Sound")]), RottenSumDiameterIntersectSquared_cmsq = sum(DiameterIntersectSquared_cmsq[which(DecayType == "Rotten")]))

# assign 1 to NULL slope factor values
cwd_plot_data$SlopeFactor[which(is.na(cwd_plot_data$SlopeFactor))] <- 1

# assign 0 to NULL sound sum diameter squared values
cwd_plot_data$SoundSumDiameterIntersectSquared_cmsq[which(is.na(cwd_plot_data$SoundSumDiameterIntersectSquared_cmsq))] <- 0

# assign 0 to NULL rotten sum diameter squared values
cwd_plot_data$RottenSumDiameterIntersectSquared_cmsq[which(is.na(cwd_plot_data$RottenSumDiameterIntersectSquared_cmsq))] <- 0

# calculate sound cwd plot biomass for each plot
cwd_plot_data$SoundCWDBiomass_MgAcre <- pmap_dbl(list(cwd_plot_data$SoundSumDiameterIntersectSquared_cmsq, "Sound", cwd_plot_data$SlopeFactor), calc_cwd_biomass)

# calculate rotten cwd plot biomass for each plot
cwd_plot_data$RottenCWDBiomass_MgAcre <- pmap_dbl(list(cwd_plot_data$RottenSumDiameterIntersectSquared_cmsq, "Rotten", cwd_plot_data$SlopeFactor), calc_cwd_biomass)

# remove extra info from cwd plot data
cwd_plot_data_trim <- cwd_plot_data[, c(1, 11, 12)]

# join tree and cwd biomass data
biomass_data <- full_join(tree_plot_data, cwd_plot_data_trim, by = "PlotID")

# add plot info for SUM503 as it was lost in the join due to no tree entries present
biomass_data[which(biomass_data$PlotID == "SUM503"), 2:9] <- c(cwd_plot_data[which(cwd_plot_data$PlotID == "SUM503"), 2:7], 0, 0)

# assign 0 to NULL sound and rotten biomass values
biomass_data$SoundCWDBiomass_MgAcre[which(is.na(biomass_data$SoundCWDBiomass_MgAcre))] <- 0
biomass_data$RottenCWDBiomass_MgAcre[which(is.na(biomass_data$RottenCWDBiomass_MgAcre))] <- 0

# join biomass and slope data
biomass_data <- left_join(biomass_data, slope_data, by = "PlotID")

# horizontal area for each plot
area_acre = (pi*14.2^2)/4046.856 # roughly 1/6 acre (0.156)

# calculate slope area for each plot
biomass_data$SlopeArea_acre <- pmap_dbl(list(area_acre, biomass_data$slope), calc_slope_area)

# calculate live tree biomass in metric tons per acre
biomass_data$LiveTreeBiomass_MgAcre <- (biomass_data$LiveTreeBiomass_kg/1000) / biomass_data$SlopeArea_acre

# calculate dead tree biomass in metric tons per acre
biomass_data$DeadTreeBiomass_MgAcre <- (biomass_data$DeadTreeBiomass_kg/1000) / biomass_data$SlopeArea_acre

# calculate BA per acre
biomass_data$BA_m2acre <- biomass_data$BA / biomass_data$SlopeArea_acre

# organzie biomass plot data
biomass_plot_data <- biomass_data[ , c(1:7, 13:18, 11:12)]

# calculate total biomass in metric tons per acre
# FIND OUT IF CWD BIOMASS IS IN METRIC OR US TONS
biomass_plot_data$TotalBiomass_MgAcre <- biomass_plot_data$LiveTreeBiomass_MgAcre + biomass_plot_data$DeadTreeBiomass_MgAcre + biomass_plot_data$SoundCWDBiomass_MgAcre + biomass_plot_data$RottenCWDBiomass_MgAcre
```

4. Potential Annual Heat Load

```{r}
# extract relevant variables
pahl_data <- biomass_plot_data[c('PlotID', 'Northing', 'Easting', 'aspect', 'slope')]

# convert UTM to LatLon
coordinates(pahl_data) <- c('Easting', 'Northing')
proj4string(pahl_data) <- CRS('+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs')

pahl_data <- spTransform(pahl_data, CRS('+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs'))

pahl_data <- as.data.frame(pahl_data)
names(pahl_data)[4:5] <- c('Lon', 'Lat')

# calculate potential annual heat load
pahl_data$folded_asp <- fold_aspect(pahl_data$aspect)

pahl_data$pahl <- pahl(pahl_data$Lat, pahl_data$slope, pahl_data$folded_asp)

# add pahl data to biomass_plot_data

biomass_plot_data <- left_join(biomass_plot_data, pahl_data[c('PlotID', 'pahl')], by = 'PlotID')
```


4.5 Make some names shorter
```{r}
names(biomass_plot_data)

colnames(biomass_plot_data)[colnames(biomass_plot_data) == "LiveTreeBiomass_MgAcre"] <- "LiveTree_MgAcre"
colnames(biomass_plot_data)[colnames(biomass_plot_data) == "DeadTreeBiomass_MgAcre"] <- "DeadTree_MgAcre"
colnames(biomass_plot_data)[colnames(biomass_plot_data) == "SoundCWDBiomass_MgAcre"] <- "SoundCWD_MgAcre"
colnames(biomass_plot_data)[colnames(biomass_plot_data) == "RottenCWDBiomass_MgAcre"] <- "RottenCWD_MgAcre"


```


5. Export to csv and shapefile.

```{r}
write.csv(tree_data, "../../data/saved/tree_data_with_biomass_all.csv")
write.csv(biomass_plot_data, "../../data/saved/biomass_data_all.csv")

#st_write(plotshp, "../../plots/biomass_allplots.shp", driver = "ESRI shapefile")
```

